Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-11-24 tesis_hiram/Chapter4/chapter4.tex
correccion

	The fit is applied by minimizing the $-\log{L}$ (NLL) with respect to the parameter $\mu$ and $\alpha_k$. The minimization is performed by using programming language ROOT. ROOT is a scientific software toolkit, based on the C++ language, which main objective is the handling of data processing, statistical analysis, including visualization and storage. The package ROOFIT is used here, because of its implementations of modeling probability distributions and statistical tools\cite{roofit}. Figure \ref{simple} and Table \ref{table1}, shows the results of the model fitting using the Asimov data for both models.

@@

 Table \ref{parameters}, shows the $\mu$ and $\alpha$ parameters. For the pre-fit, values of $\alpha$ are set to zero and $\mu$ is set to 1. After the fit, $\alpha$ parameters set to zero indicates that in the fit, the values of the backgrounds didn't change for the SM and the $\mu$ parameter set to 1 indicates that the signal strength also didn't change.

@@

		\begin{figure}[ht]
		\centering
		\begin{minipage}[b]{0.48\textwidth}
			\includegraphics[width=\textwidth]{Chapter4/simple.png}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.48\textwidth}
			\includegraphics[width=\textwidth]{Chapter4/simple-kt-1.png}
		\end{minipage}
		\caption{Post fit signal and background yields for tH process for SM (Left) and $k_t=-1$ (Right).
			%In the box below each distribution, the ratio of the observed and predicted event yields is shown
		}
		\label{simple}
	\end{figure}
	\begin{table}[ht]
		\centering
		\caption{Postfit yields for the fit to the Asimov data corresponding to 35.9 fb$^{-1}$. The uncertainty given is the combined statistical plus systematic.}
		\begin{tabular}{ccc}
			\hline
			Process & SM & $k_{t}$=-1 \\
			\hline
			$t\bar{t}W$ & 68$\pm$8.9& 68 $\pm$8.9 \\
			$t\bar{t}Z$ & 25.9$\pm$3.8&25.9$\pm$3.8\\
			$WZ$ & 15.1$\pm$7.4& 15.1$\pm$7.4\\
			Rares & 20.8$\pm$4.8& 20.8$\pm$4.8 \\
			Fakes & 80.9$\pm$9.0& 80.9$\pm$8.9 \\
			$t\bar{t}H$ & 24.2$\pm$2.0 & 24.2$\pm$2.0 \\
			\hline
			$tH$& 2.1$\pm$16.5 &26.2$\pm$13.1 
		\end{tabular}
		\label{table1}
	\end{table}
	\begin{table}[ht!]
		\small
		\centering
		\caption{$\alpha$ and $\mu$ values for the fit to Asimov data corresponding to 35.9 fb $^{-1}$}
		\begin{tabular}{ccc}
			\hline
			Parameter & SM &$k_t$=-1\\
			\hline
			$\mu$ & 1.00 $\pm$ 7.74& 1.0 $\pm$ 0.5\\
			$\alpha_{t\bar{t}W}$& 0.00 $\pm$ 0.89& 0.00 $\pm$ 089\\
			$\alpha_{t\bar{t}Z}$ & 0.00 $\pm$ 0.98& 0.00 $\pm$ 0.98\\
			$\alpha_{WZ}$ & 0.00 $\pm$ 0.97& 0.00 $\pm$ 0.97\\
			$\alpha_{Rares}$ &0.00 $\pm$ 0.95&0.00 $\pm$ 0.98 \\
			$\alpha_{Fakes}$ & 0.00 $\pm$ 0.96& 0.00 $\pm$ 0.95\\
			$\alpha_{t\bar{t}H}$ &0.00 $\pm$ 0.98& 0.00 $\pm$ 0.98\\
		\end{tabular}
		\label{parameters}
	\end{table}

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[scale=0.5]{Chapter4/chi}
		\caption{Illustration of the $\chi^2$ distribution %with 7 degrees of freedom 
			used for the limit estimation}
		\label{chi}
	\end{figure}

In order to find the upper limit value of $\mu$, we search for the largest value of $\mu$ such that $q$ remains below the shaded region shown in figure \ref{chi}, where $\alpha$=0.05 in this case. It is useful to scan $q$ as function of $\mu$, which is normally a parabolic function as shown in figure \ref{scanl}. In the SM case, the scan has a wider shape in comparison to the $k_t$=-1. This is due to larger statistical uncertainty on the $\mu$ parameter. 
	Using the above technique, we find an expected upper limit $\mu < $17.0 for SM model and $\mu < $ 2.3 for $k_t$=-1 model. 
		\begin{figure}[ht]
		\centering
		\begin{minipage}[b]{0.49\textwidth}
			\includegraphics[width=\textwidth,height=0.31\textheight]{Chapter4/Likelihood.png}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.49\textwidth}
			\includegraphics[width=\textwidth,height=0.31\textheight]{Chapter4/Likelihood-kt-1.png}
		\end{minipage}
		\caption{Likelihood scan for SM (Left) and $k_t$=-1 (Right).}
		\label{scanl}
	\end{figure}
	\pagebreak
	%%sigma son las probabilidades que corresponden a 1sigma=68, 2sigma=95, 3sigma=98 y sigma son los limites de la integral de una gaussiana
	\section{Extrapolation to higher luminosities}
	In the past section, it was shown that the signal strength with an integrated luminosity of 35.9 fb$^{-1}$ is not significant. The LHC has already collected data for integrated luminosity of 150 fb$^{-1}$ for run 2 which finished in 2018. Run 3 starts in 2021 and is expected an integrated luminosity of 300fb$^{-1}$ and for the high luminosity phase of LHC, expected to collect 3000 fb$^{-1}$. Therefore, it is interesting to study the sensitivity of the signal strength at higher luminosities. 
	An extrapolation (scale) is applied to the Asimov data and the statistical uncertainty is adjusted accordingly. For this analysis we assume the systematic uncertainties assigned to the background are the same.

	In figure \ref{hm}, an estimated prediction of the signal and background events for higher luminosities is shown. As the luminosity increases, the number of events for both signal and backgrounds also increases, as previously explained in Chapter 1. But even with the increase of number of events, the statistical uncertainty of the data is too large with respect to the expected signal. In the likelihood scan for higher luminosities, as the luminosity increases, the relative uncertainty on the signal strength $\mu$ decreases. This is caused by the increased events available for the fit.

@@

		\begin{table}[ht]
		\caption{Results of $\mu$ and upper limits for Asimov extrapolations for SM  and $k_t$=-1 models at 95$\%$ confidence level.}
		\begin{tabular}{|c|c|c|c|c|}
			\hline
			\multirow{2}{*}{Luminosity (fb$^{-1}$)}  &\multicolumn{2}{c|}{SM ($k_t$=1)} &\multicolumn{2}{c|}{$k_t$=-1 scenario}\\
			\hhline{~----}	&$\mu$  &$\mu$ upper limit & $\mu$ &$\mu$ upper limit or interval\\
			\hline
			35.9 & 1.0 $\pm$ 7.7 & 17 & 1.0 $\pm$ 0.5 & 2.3 \\
			\hline
			150& 1.0 $\pm$ 6.7& 11 & 1.0 $\pm$ 0.4 &1.8\\
			\hline
			300&1.0 $\pm$ 4.3 &8.7 & 1.0 $\pm$ 0.3 &1.5 \\
			\hline
			3000&1.0 $\pm$ 1.7 & 4.3 &	 1.0 $\pm$ 0.1 & 0.9-1.1\\
			\hline
		\end{tabular}
		\label{upper}
	\end{table}

	However, the uncertainty of $\mu$ for the SM case is 170$\%$ for  in the 3000 fb$^{-1}$ case, making impossible to get an evidence of a Higgs boson in the $tH$ process for the same sign muon channel. Even at higher luminosities, the uncertainty of $\mu$ in SM case is still very big.
	In the $k_t$=-1 model the uncertainty of $\mu$ decreases with higher luminosities giving better results. This is caused by the fact that the number of events for the $tH$ signal is greater. For 3000 fb$^{-1}$ the uncertainty on $\mu$ is $~10\%$, and it has an upper limit that rounds between 0.9-1.1. In the likelihood scan, the statistical test $q$ gets very high values at $\mu$>1.1. This indicates that for large integrated luminosities this model can be fully ruled out. This can be visualized in the likelihood scan of  3000 fb$^{-1}$, where the curve is very closed and values of $q$ are very high, giving incompatibility of the model with data.

>>>>

	The fit is applied by minimizing the negative logarithm of the likelihood function (NLL) with respect to the parameter $\mu$ and $\alpha_k$. The minimization is performed by using programming language ROOT. ROOT is a scientific software toolkit, based on the C++ language, which main objective is the handling of data processing, statistical analysis, including visualization and storage. The package ROOFIT is used here, because of its implementations of modeling probability distributions and statistical tools\cite{roofit}. Figure \ref{simple} and Table \ref{table1}, shows the results of the fit to the Asimov data.

@@

 Table \ref{parameters}, shows the $\mu$ and $\alpha$ parameters. For the pre-fit, values of $\alpha$ are set to zero and $\mu$ is set to 1. After the fit, $\alpha$ parameters set to zero indicates that in the fit, the values of the backgrounds did not change for the SM and the $\mu$ parameter set to 1 indicates that the signal strength also did not change.

@@

		\begin{figure}[ht]
		\centering
		\begin{minipage}[b]{0.48\textwidth}
			\includegraphics[width=\textwidth]{Chapter4/simple.png}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.48\textwidth}
			\includegraphics[width=\textwidth]{Chapter4/simple-kt-1.png}
		\end{minipage}
		\caption[Post-fit Signal and Backgrounds for $tH$ process]{Post-fit signal and background yields for $tH$ process for SM (Left) and $k_t=-1$ (Right).
			%In the box below each distribution, the ratio of the observed and predicted event yields is shown
		}
		\label{simple}
	\end{figure}
	\begin{table}[ht]
		\centering
		\caption[Postfit yields for the fit to the Asimov data corresponding to 35.9 fb$^{-1}$]{Postfit yields for the fit to the Asimov data corresponding to 35.9 fb$^{-1}$. The uncertainty given is the combined statistical plus systematic.}
		\begin{tabular}{ccc}
			\hline
			Process & SM & $k_{t}$=-1 \\
			\hline
			$t\bar{t}W$ & 68$\pm$8.9& 68 $\pm$8.9 \\
			$t\bar{t}Z$ & 25.9$\pm$3.8&25.9$\pm$3.8\\
			$WZ$ & 15.1$\pm$7.4& 15.1$\pm$7.4\\
			Rares & 20.8$\pm$4.8& 20.8$\pm$4.8 \\
			Fakes & 80.9$\pm$9.0& 80.9$\pm$8.9 \\
			$t\bar{t}H$ & 24.2$\pm$2.0 & 24.2$\pm$2.0 \\
			\hline
			$tH$& 2.1$\pm$16.5 &26.2$\pm$13.1 
		\end{tabular}
		\label{table1}
	\end{table}
	\begin{table}[ht!]
		\small
		\centering
		\caption{$\alpha$ and $\mu$ values for the fit to Asimov data corresponding to 35.9 fb $^{-1}$}
		\begin{tabular}{ccc}
			\hline
			Parameter & SM &$k_t$=-1\\
			\hline
			$\mu$ & 1.00 $\pm$ 7.74& 1.0 $\pm$ 0.5\\
			$\alpha_{t\bar{t}W}$& 0.00 $\pm$ 0.89& 0.00 $\pm$ 089\\
			$\alpha_{t\bar{t}Z}$ & 0.00 $\pm$ 0.98& 0.00 $\pm$ 0.98\\
			$\alpha_{WZ}$ & 0.00 $\pm$ 0.97& 0.00 $\pm$ 0.97\\
			$\alpha_{Rares}$ &0.00 $\pm$ 0.95&0.00 $\pm$ 0.98 \\
			$\alpha_{Fakes}$ & 0.00 $\pm$ 0.96& 0.00 $\pm$ 0.95\\
			$\alpha_{t\bar{t}H}$ &0.00 $\pm$ 0.98& 0.00 $\pm$ 0.98\\
		\end{tabular}
		\label{parameters}
	\end{table}

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[scale=0.5]{Chapter4/chi}
		\caption[$\chi^2$ Distribution]{Illustration of the $\chi^2$ distribution with 7 degrees of freedom used for the limit estimation}
		\label{chi}
	\end{figure}

In order to find the upper limit value of $\mu$, we search for the largest value of $\mu$ such that $q$ remains below the shaded region shown in figure \ref{chi}, where $\alpha$=0.05 in this case. It is useful to scan $q$ as function of $\mu$, which is normally a parabolic function as shown in figure \ref{scanl}. In the SM case, the scan has a wider shape in comparison to the $k_t$=-1. This is due to larger statistical uncertainty on the $\mu$ parameter. 
	Using the above technique, we find an expected upper limit $\mu < $17.0 for SM model and $\mu < $ 2.3 for $k_t$=-1 model. 
		\begin{figure}[ht]
		\centering
		\begin{minipage}[b]{0.49\textwidth}
			\includegraphics[width=\textwidth,height=0.31\textheight]{Chapter4/Likelihood.png}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.49\textwidth}
			\includegraphics[width=\textwidth,height=0.31\textheight]{Chapter4/Likelihood-kt-1.png}
		\end{minipage}
		\caption[Likelihood scan for SM and $k_t$=-1]{Likelihood scan for SM (Left) and $k_t$=-1 (Right).}
		\label{scanl}
	\end{figure}
	\pagebreak
	%%sigma son las probabilidades que corresponden a 1sigma=68, 2sigma=95, 3sigma=98 y sigma son los limites de la integral de una gaussiana
	\section{Extrapolation to higher luminosity}
	In the past section, it was shown that the signal strength with an integrated luminosity of 35.9 fb$^{-1}$ is not significant. The LHC has already collected data for integrated luminosity of 150 fb$^{-1}$ for run 2 which finished in 2018. Run 3 starts in 2021 and is expected an integrated luminosity of 300fb$^{-1}$ and for the high luminosity phase of LHC, expected to collect 3000 fb$^{-1}$. Therefore, it is interesting to study the sensitivity of the signal strength at higher luminosity. 
	An extrapolation (scale) is applied to the Asimov data and the statistical uncertainty is adjusted accordingly. For this analysis we assume the systematic uncertainties assigned to the background are the same.

	In figure \ref{hm}, an estimated prediction of the signal and background events for higher luminosity is shown. As the luminosity increases, the number of events for both signal and backgrounds also increases, as previously explained in Chapter 1. But even with the increase of number of events, the statistical uncertainty of the data is too large with respect to the expected signal. In the likelihood scan for higher luminosity, as the luminosity increases, the relative uncertainty on the signal strength $\mu$ decreases. This is caused by the increased events available for the fit.

@@

		\begin{table}[ht]
		\caption[Results of $\mu$ and upper limits for SM and $k_t$=-1]{Results of $\mu$ and upper limits for Asimov extrapolations for SM  and $k_t$=-1 models at 95$\%$ confidence level.}
		\begin{tabular}{|c|c|c|c|c|}
			\hline
			\multirow{2}{*}{Luminosity (fb$^{-1}$)}  &\multicolumn{2}{c|}{SM ($k_t$=1)} &\multicolumn{2}{c|}{$k_t$=-1 scenario}\\
			\hhline{~----}	&$\mu$  &$\mu$ upper limit & $\mu$ &$\mu$ upper limit or interval\\
			\hline
			35.9 & 1.0 $\pm$ 7.7 & 17 & 1.0 $\pm$ 0.5 & 2.3 \\
			\hline
			150& 1.0 $\pm$ 6.7& 11 & 1.0 $\pm$ 0.4 &1.8\\
			\hline
			300&1.0 $\pm$ 4.3 &8.7 & 1.0 $\pm$ 0.3 &1.5 \\
			\hline
			3000&1.0 $\pm$ 1.7 & 4.3 &	 1.0 $\pm$ 0.1 & 0.9-1.1\\
			\hline
		\end{tabular}
		\label{upper}
	\end{table}

	However, the uncertainty of $\mu$ for the SM case is 170$\%$ for  in the 3000 fb$^{-1}$ case, making impossible to get an evidence of a Higgs boson in the $tH$ process for the same sign muon channel. Even at higher luminosity, the uncertainty of $\mu$ in SM case is still very big.
	In the $k_t$=-1 model the uncertainty of $\mu$ decreases with higher luminosity giving better results. This is caused by the fact that the number of events for the $tH$ signal is greater. For 3000 fb$^{-1}$ the uncertainty on $\mu$ is $~10\%$, and it has an upper limit that rounds between 0.9-1.1. In the likelihood scan, the statistical test $q$ gets very high values at $\mu$>1.1. This indicates that for large luminosity this model can be fully ruled out. This can be visualized in the likelihood scan of  3000 fb$^{-1}$, where the curve is very closed and values of $q$ are very high, giving incompatibility of the model with data.

### 2019-11-24 tesis_hiram/Chapter5/chapter5.tex
correccion

\chapter{Conclusion and outlook}
In this work, we explored the sensitivity of the production process of Higgs plus single top.
The $tH$ process is an interesting channel in proton-proton collisions in the LHC. This process has not been observed experimentally. Recently CMS published a search for $tH$ production using 35.9 fb$^{-1}$ from Run 2 . Using information from this dataset, we studied the expected sensitivity using an Asimov data and systematic uncertainties for the final state with same sign pairs of muons.

Using this dataset for 35.9 fb$^{-1}$, the uncertainty for the $tH$ signal strength $\mu$ is around eight times than the value of $\mu$ for the SM. In this case, we estimated an upper limit on $\mu$ of 17 at 95$\%$ confidence level. 
For the inverted coupling scenario ($k_t$ =-1), the uncertainty after the fit was around 50$\%$, this is due to the larger expected number of signal events

We also studied the sensitivity for larger integrated luminosities expected in future runs of the LHC. We find that for the SM scenario the expected uncertainty even at the largest luminosities is not enough to observe the signal and only an upper limit on $\mu$ is placed at 4.3. In the case of the inverted coupling scenario, a possible signal can be observed with an uncertainty of 10$\%$ at 3000 fb$^{-1}$.

>>>>

\chapter{Conclusion and outlook}
In this work, we explored the sensitivity of the production process of Higgs plus single top.
The $tH$ process is an interesting channel in proton-proton collisions in the LHC. This process has not been observed experimentally. Recently CMS published a search for $tH$ production using 35.9 fb$^{-1}$ from Run 2\cite{th1}. Using information from this publication, we studied the expected sensitivity using an Asimov data and systematic uncertainties for the final state with same sign pairs of muons.

For an integrated luminosity of 35.9 fb$^{-1}$, we obtained an expected uncertainty for the $tH$ signal strength $\mu$ is around eight times than the value of $\mu$ for the SM. In this case, we estimated an upper limit on $\mu$ of 17 at 95$\%$ confidence level. 
For the inverted coupling scenario ($k_t$ =-1), the uncertainty after the fit was around 50$\%$, this is due to the larger expected number of signal events.

We also studied the sensitivity for larger integrated luminosity expected in future runs of the LHC. We find that for the SM scenario the expected uncertainty even at the largest integrated luminosity is not enough to observe the signal and only an upper limit on $\mu$ is placed at 4.3. In the case of the inverted coupling scenario, a possible signal can be observed with an uncertainty of 10$\%$ at 3000 fb$^{-1}$.
